---
title: "data cleaning and linear regression"
author: Julia
output: pdf_document
---

```{r}
library(tidyverse)
library(ggplot2)
```

# original input file with information on public companies
```{r}
ipo<-read.csv("basicIPOs.csv")
```

# generate a random dataset for further testing samples
```{r}
random_ipo= ipo[sample(1:nrow(ipo)), ]
random_ipo$region<-state.name[match(random_ipo$state,state.abb)]
```
```{r}
write.csv(random_ipo,"ipo_random.csv", row.names = FALSE)
```


# first 500 companies from Diffbot
```{r}
temp_df<-read.csv("selected500.csv")
cominfo1<-read.csv("500compinfo.csv")
cominfo1$Industry=temp_df$Industries
```

#second 1500 companies from Diffbot
```{r}
cominfo2<-read.csv("500to1500compinfo.csv")
```

# another 2000 companies from Diffbot
```{r}
cominfo3<-read.csv("compinfo2000.csv")
```

#combine all companies
```{r}
cominfo <- rbind(cominfo1, cominfo2,cominfo3)
```

# change empty cells to NAs for further selecting and making changes
```{r}
cominfo <- cominfo %>% mutate_all(na_if,"")
```


# divide the dataset into VC companies and Non VC companies
```{r}
vc<-cominfo%>%filter(VC_dummy=="1")
nonvc<-cominfo%>%filter(VC_dummy=="0")
```

# tibble listing the number of VC backed companies in each industry  
```{r}
vc_noin<-vc%>%drop_na(c("Industry"))
vcIndustry<-vc_noin %>%
  group_by(Industry)%>%summarise(Freq=n())%>%arrange(desc(Freq))
```

```{r}
vcIndustry
```

# tibble listing the number of NonVC companies in each industry
```{r}
nonvc_noin<-nonvc%>%drop_na(c("Industry"))
nonvcIndustry<-nonvc_noin %>%
  group_by(Industry)%>%summarise(Freq=n())%>%arrange(desc(Freq))
```
```{r}
nonvcIndustry
```

```{r}
df1<-merge(nonvcIndustry,vcIndustry,by="Industry")
df1$total<-df1$Freq.x+df1$Freq.y
df1
```

# make a ggplot of the types of industries invested by VC of the 20 most common industries
```{r}
df1_vc<-head(df1%>%arrange(desc(Freq.y)),20)
```

```{r}
ggplot(df1_vc) + geom_bar(aes(x=Industry, y=total), stat="identity",fill="pink") +  geom_bar(aes(x=Industry   , y=Freq.y, fill = "VC"), stat="identity",fill="blue")+ coord_flip() 
```
# make a ggplot of the types of industries not invested by VC of the 20 most commonly not invested industries
```{r}
df1_nonvc<-head(df1%>%arrange(desc(Freq.x)),20)
```
```{r}
ggplot(df1_nonvc) + geom_bar(aes(x=Industry, y=total), stat="identity",fill="pink") +  geom_bar(aes(x=Industry , y=Freq.x, fill = "NONVC"), stat="identity",fill="orange")+ coord_flip() 
```


# linear model with number of employees as the y variable and only VC_dummy as the x variable 
```{r}
model_emp1<-lm(nbEmployees~VC_dummy,cominfo)
summary(model_emp1)
```


```{r}
library(lmtest)
library(sandwich)
```

# linear regression model with y variable as the number of employees and x variables are VC_dummy, industry, age, and state

```{r}
model_emp2<-lm(nbEmployees~VC_dummy+factor(Industry)+age+factor(region),cominfo)
summary(model_emp2)
```



```{r}
extracted_col = fullcom["nbLocations"]
cominfo$nbLocations = extracted_col
```

# linear model with number of locations as the y variable and VC, industry, age, and state as x variable after log transformation
```{r}
model_loc<-lm(log(unlist(nbLocations))~VC_dummy+factor(Industry)+log(age)+factor(region),cominfo)
summary(model_loc)
```


# # linear regression model with y variable as the number of employees and x variables are VC_dummy, industry, age, and state after log transformation
```{r}
l.cominfo<-cominfo%>%mutate(log.Employees=log(nbEmployees),log.age=log(age))%>%dplyr::select(log.Employees,VC_dummy,Industry,log.age,age,region)
l.cominfo[is.na(l.cominfo) | l.cominfo=="-Inf"] = NA
l.model_emp=lm(log.Employees~VC_dummy+factor(Industry)+age+factor(region),l.cominfo)
```

```{r}
summary(l.model_emp)
```

# linear regression model with y variable as the number of patents each company has between the year they are founded and the year they went public and x variables are VC_dummy, industry, age, and state before and after log tranformation
```{r}
cominfo$age=2022-cominfo$Year_Founded_Patton
l.cominfo<-cominfo%>%mutate(log.patent=log(patent_tillipo),log.age=log(age))%>%dplyr::select(patent_tillipo,log.patent,VC_dummy,Industry,log.age,age,region)
l.cominfo[is.na(l.cominfo) | l.cominfo=="-Inf"] = NA
```

# before:
```{r}
model_patent<-lm(patent_tillipo~VC_dummy+factor(Industry)+age+factor(region),cominfo)
```

```{r}
summary(model_patent)
```

# after:
```{r}
l.model_patent<-lm(patent_tillipo~VC_dummy+factor(Industry)+log.age+factor(region),l.cominfo)
```

```{r}
summary(l.model_patent)
```

